---
title: "TO 628 Final Project: Spotify Hits Data"
author: "Team 8: MBAi"
date: "4/9/2023"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: readable
    highlight: tango
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intermediate Deliverable

## Link to GitHub repo

https://github.com/sammysunnyshine/TO-628-Final-Project.git

## Data Set Details

**The Spotify Hit Predictor Dataset: 2000-2009**

This data set includes songs on Spotify that were produced between Jan 01, 2000 and Dec 31, 2009. Each song has been scored across a number of variables; these scores are presented along with other details about the songs, such as duration, tempo, etc.

Variables:

- track: The Name of the track.

- artist: The Name of the Artist.

- uri: The resource identifier for the track.

- danceability: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable. 

- energy: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy. 

- key: The estimated overall key of the track. Integers map to pitches using standard Pitch Class notation. E.g. 0 = C, 1 = C?/D?, 2 = D, and so on. If no key was detected, the value is -1.

- loudness: The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typical range between -60 and 0 db. 

- mode: Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.

- speechiness: Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks. 

- acousticness: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic. The distribution of values for this feature look like this:

- instrumentalness: Predicts whether a track contains no vocals. “Ooh” and “aah” sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly “vocal”. The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content. Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0. The distribution of values for this feature look like this:

- liveness: Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live.

- valence: A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).

- tempo: The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration. 

- duration_ms:  The duration of the track in milliseconds.

- time_signature: An estimated overall time signature of a track. The time signature (meter) is a notational convention to specify how many beats are in each bar (or measure).

- chorus_hit: This the the author's best estimate of when the chorus would start for the track. Its the timestamp of the start of the third section of the track. This feature was extracted from the data received by the API call for Audio Analysis of that particular track.

- sections: The number of sections the particular track has. This feature was extracted from the data received by the API call for Audio Analysis of that particular track.

- target: The target variable for the track. It can be either '0' or '1'. '1' implies that this song has featured in the weekly list (Issued by Billboards) of Hot-100 tracks in that decade at least once and is therefore a 'hit'. '0' Implies that the track is a 'flop'.

## Business Question

What makes a song turn into a "hit" or a "flop"? Our goal is to build a model that predicts which songs will be successful on Spotify, which serves as a proxy for the overall music market.

This will be extremely useful to artists, managers, producers, and all other parties who have stakes in launching new music. Our model will help identify what characteristics make a song successful, regardless of genre, and help evaluate existing songs and new songs in production. Why does this matter?

- Maximizing ROI: creators, their talent managers, and music labels can better prioritize to which songs they should allocate more resources, e.g. which songs to record music videos for and include in an album 

- Artist development: music labels and talent managers can identify and develop emerging talent if they have early indications that their songs have a high potential to become hits

- Marketing strategies: Many artists are savvy marketers and strategically order and time song releases (e.g. Taylor Swift is well-known for that). A good prediction model could supplement their instincts and business acumen in their decision-making

- Streaming services and payment models: Given there have been ongoing discussions in the news cycle around payment models used by streaming services like Spotify and Apple Music, with the latter being criticized for inadequately compensating creators, prediction models for song's success in the market could help open the door for alternative payment arrangements. Currently, a song could easily take $1K to produce, yet 100K streams could results in approx. $400 or less of a paycheck. 

*Our model will be able to determine with a high level of confidence whether a song will be a "hit" or a "flop."* Being able to predict if a song would be a hit can help stakeholders in the music industry make data-driven decisions, optimize their investments into new music and music videos, and ultimately increase revenue.

*Sources:* 
https://www.usatoday.com/story/life/2022/10/22/how-much-per-spotify-stream/8094437001/
https://dittomusic.com/en/blog/how-much-does-spotify-pay-per-stream/

## Read and Clean Data
```{r}
spot <- read.csv("spotify_hits.csv")

spot$uri <- NULL

# Delete the tracks field because it's essentially individual IDs
spot$track <- NULL

# Deleting artist - too many unique values
spot$artist <- NULL

# There are no NA values
# There is no need to create factors

str(spot)
summary(spot)

```

See code comments for details on cleaning. Dummy variables are not needed.

## Randomize the Data Set
```{r}
spot_random <- spot[sample(1:nrow(spot)), ]
```

## Normalize the Data
```{r}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

spot_norm <- as.data.frame(lapply(spot_random, normalize))
```

## Split Test and Train
```{r}
set.seed(13090890)
test_set <- sample(1:nrow(spot_norm), 0.5*nrow(spot_norm))
spot_test <- spot_norm[test_set, ]
spot_train <- spot_norm[-test_set, ]

```

Created spot_test and spot_train with a 50/50 random sample.

## Build Regression Model
```{r}
model_lr <- glm(target ~ ., family = "binomial", data = spot_train)
summary(model_lr)

```

Seven variables are significant: 

- Danceability
- Energy
- Loudness
- Mode
- Acousticness
- Instramentalness
- Time_signature

## Predict Test Data
```{r}
spot_pred <- predict(model_lr, spot_test, type = "response")

summary(spot_pred)

spot_bin_pred <- ifelse(spot_pred >= 0.5, 1, 0) # Given the data is relatively balanced, we use 50% threshold

table(spot_bin_pred)

```
Based on a 50% prediction threshold, the model predicts that of the `r 1341 + 1595` observations in the test data set, approximately `r round(100*1595/(1341+1595),0)` percent will be hits. 

## Evaluate Model
```{r}
library(caret)
confusionMatrix(as.factor(spot_bin_pred), as.factor(spot_test$target), positive = "1")

```

The model has an accuracy of 82% and a Kappa of 63%.

# Final Deliverable

## Flag and Split Popular Artists
```{r}
spot2 <- read.csv("spotify_hits.csv")
criteria <- read.csv("popular_artists.csv", header = FALSE)
criteria_vec <- as.character(criteria$V1)

# Explore top artists
head(sort(table(spot2$artist), decreasing = TRUE), 20)

spot2$uri <- NULL

# Delete the tracks field because it's essentially individual IDs
spot2$track <- NULL

# Creating dummy variable to distinguish popular and lesser-known artists
spot2$artist_popularity <- ifelse(spot2$artist %in% criteria_vec, 1, 0)

# Deleting artist - too many unique values
spot2$artist <- NULL

# Split Popular and Lesser-Known Artists
spot_unknown <- spot2[spot2$artist_popularity == 0, ]
spot_popular <- spot2[spot2$artist_popularity == 1, ]

spot_unknown$artist_popularity <- NULL
spot_popular$artist_popularity <- NULL 

str(spot_unknown)
summary(spot_unknown)

```

Popular artists are popular for a variety of reasons that are entirely separate from this data set -- marketing, exposure, appearance, connections, etc. -- that it does not make sense to run a predictive model on the data we have. However, among *unknown* artists who have limited existing support, it is far more important to predict who will hit and who will flop. This group of artists has the highest volatility but potentially the highest payoff, so we will focus here.

## Randomize the Unknown Artists
```{r}
spot_uk_random <- spot_unknown[sample(1:nrow(spot_unknown)), ]
```

## Normalize the Data
```{r}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

spot_uk_norm <- as.data.frame(lapply(spot_uk_random, normalize))

summary(spot_uk_norm)
```

## Split Test and Train
```{r}
set.seed(13090890)
test_set_uk <- sample(1:nrow(spot_uk_norm), 0.5*nrow(spot_uk_norm))
spot_test_uk <- spot_uk_norm[test_set_uk, ]
spot_train_uk <- spot_uk_norm[-test_set_uk, ]

summary(spot_test_uk)

spot_test_uk_nl <- spot_test_uk[ , -match("target", names(spot_uk_norm))]
spot_train_uk_nl <- spot_train_uk[ , -match("target", names(spot_uk_norm))]

# Now the response (aka Labels) - only the target column
spot_test_uk_labels <- spot_test_uk[, "target"]
spot_train_uk_labels <- spot_train_uk[ , "target"]

```

Created spot_test_uk and spot_train_uk with a 50/50 random sample.

## Build Regression Model
```{r}
model_lr_uk <- glm(target ~ ., family = "binomial", data = spot_train_uk)
summary(model_lr_uk)

```

Mostly the same variables are significant for unknown artists, but instead of Time_signature, Liveness is significant.

- Danceability
- Energy
- Loudness
- Mode
- Speechiness
- Acousticness
- Instrumentalness
- Liveness

### Predict Test Data
```{r}
pred_lr_uk <- predict(model_lr_uk, spot_test_uk, type = "response")

summary(pred_lr_uk)

bin_pred_lr_uk <- ifelse(pred_lr_uk >= 0.5, 1, 0) # Given the data is relatively balanced, we use 50% threshold

table(bin_pred_lr_uk)

```

Based on a 50% prediction threshold, the model predicts that of the `r 1379 + 1041` observations in the test data set, approximately `r round(100*1041/(1379 + 1041),0)` percent will be hits. 

### Evaluate Model
```{r}
library(caret)
confusionMatrix(as.factor(bin_pred_lr_uk), as.factor(spot_test_uk$target), positive = "1")

```

Our initial Logistic Regression model for unknown artists is very strong: accuracy of 81% and **Kappa of 62%**.

## Build KNN Model and Predict

```{r}
library(class)

model_knn_uk <- knn(spot_train_uk_nl, spot_test_uk_nl, spot_train_uk_labels, k = 70)

```

### Evaluate model
```{r}
library(caret)

confusionMatrix(as.factor(model_knn_uk), as.factor(spot_test_uk_labels), positive = "1")

```

Lower accuracy and Kappa than LR model, but still very strong.

K = 100; Acc 77.3%, Kappa 54.9%

K = 85; Acc 77.4%, Kappa 54.9%

**K = 70; Acc 77.5%, Kappa 55.1%**

K = 50; Acc 77.4%, Kappa 54.8%

K = 35; Acc 77.6%, Kappa 55.6%

K = 20; Acc 77.8%, Kappa 55.3%

K = 10; Acc 77.2%, Kappa 53.8%

K = 5; Acc 76.4%, Kappa 52.0%



While the above numbers will change each time the model runs, they are demonstrative of the differences among K values.

We started with K = 70 (approx. square root of row count). Given all the Kappas are very similar, and we prefer a higher K to avoid overfitting, we will go with K = 70.

## Build ANN Model

```{r}
library(neuralnet)

model_ann_uk <- neuralnet(target ~ ., data = spot_train_uk, hidden = 2)
plot(model_ann_uk)

```

### Predict and Evaluate

```{r}
pred_ann_uk <- predict(model_ann_uk, spot_test_uk)
summary(pred_ann_uk)

bin_ann_uk <- ifelse(pred_ann_uk >= 0.5, 1, 0) # Given the data is relatively balanced, we use 50% threshold
library(caret)
confusionMatrix(as.factor(bin_ann_uk), as.factor(spot_test_uk$target), positive = "1")


```

Even better Kappa than Logistic Regression. While the following numbers will change each time the model runs, they are demonstrative of the differences among neuron combinations.

ANN Kappa with one neuron: 64%. ANN Kappa with two neurons: 67%. ANN Kappa with three neurons: 66%. ANN Kappa with (3, 2, 1) neurons: 63%. Other multi-layer attempts were all below 64%.

Adding additional neurons and layers does not materially and consistently improve the Kappa of this model; we ran each combination several times and got a range of Kappas between 63% and 67%, but no combination was ever consistently the best.

**Because two neurons gave us the highest individual Kappa, we will go forward with that number.**

## Build Decision Tree

```{r}
library(C50)
spot_tree_uk <- C5.0(as.factor(target) ~ ., data = spot_train_uk)

plot(spot_tree_uk)
summary(spot_tree_uk)
```

### Predict and Evaluate

```{r}
pred_tree_uk <- predict(spot_tree_uk, spot_test_uk)

library(caret)
confusionMatrix(as.factor(pred_tree_uk), as.factor(spot_test_uk$target), positive = "1")

```

Kappa still greater than 60%, but lower than ANN.

## Build Random Forest

```{r}
#install.packages("randomForest")
library(randomForest)
spot_forest_uk <- randomForest(as.factor(target) ~ ., data = spot_train_uk)

plot(spot_forest_uk)
summary(spot_forest_uk)
```

### Predict and Evaluate

```{r}
pred_forest_uk <- predict(spot_forest_uk, spot_test_uk)

library(caret)
confusionMatrix(as.factor(pred_forest_uk), as.factor(spot_test_uk$target), positive = "1")

```

**Random forest has the highest Kappa of any individual model at almost 68%.** Some trials reached as high as almost 70%.

## Combine Models

```{r}
pred_comb_uk <- data.frame(pred_lr_uk, model_knn_uk, pred_ann_uk, pred_tree_uk, pred_forest_uk, spot_test_uk$target)

str(pred_comb_uk)
summary(pred_comb_uk)
```

### New Test and Train Data Sets

```{r}
set.seed(13090890)
test_set_comb <- sample(1:nrow(pred_comb_uk), 0.3*nrow(pred_comb_uk))

spot_test_comb <- pred_comb_uk[test_set_comb, ]
spot_train_comb <- pred_comb_uk[-test_set_comb, ]

```

## Build Stacked Model

```{r}
library(C50)
spot_tree_comb <- C5.0(as.factor(spot_test_uk.target) ~ ., data = spot_train_comb)

plot(spot_tree_comb)
summary(spot_tree_comb)
```

The stacked-model decision tree is telling us the following:

If the Random Forest model predicts a song will be a hit, the stacked model will predict it will be a hit; otherwise, it will not. Nothing matters other than the random forest.

### Predict and Evaluate

```{r}
pred_tree_comb <- predict(spot_tree_comb, spot_test_comb)

library(caret)
confusionMatrix(as.factor(pred_tree_comb), as.factor(spot_test_comb$spot_test_uk.target), positive = "1")

```

The stacked model is extremely strong and has a **Kappa of 68%**, better than each individual model.

## Incorporate Cost Matrix

There are approximately the same number of false negatives and false positives in the prediction.

A false positive in this case represents songs we predict will be hits but they flop. The loss here is the amount of money invested in the production of these songs and the reputation hit from having recommended their production. Even though the average cost to produce a song is only approx. $2.5K, this is bare minimum. Add in all the marketing and other support money, and the reputational damage if a promoted song fails, and this could result in a substantial loss. We want to minimize these errors.

A false negative in this case represents songs we predict will be flops but they ultimately become hits. The loss here is an opportunity cost -- the result of a decision not taken. While a massive hit can produce a lot of money for those involved, as Sanjeev says, nobody's ever been fired for that kind of mistake. While not ideal, we'd prefer these mistakes to up-front financial losses.

We estimate a false positive is twice as damaging as a false negative.

```{r}
error_cost <- matrix(c(0, 2, 1, 0), nrow = 2)
error_cost

model_error_uk <- C5.0(as.factor(spot_test_uk.target) ~ ., data = spot_train_comb, costs = error_cost)

pred_error_uk <- predict(model_error_uk, spot_test_comb)

confusionMatrix(as.factor(pred_error_uk), as.factor(spot_test_comb$spot_test_uk.target), positive = "1")
```

The Kappa of the stacked model has been reduced slightly, to around 65-66%, but this is still very strong. However, we also ran trials in which we assumed a much heavier cost to false positives, and unfortunately the Kappa dropped substantially to around 45%. A larger the negative impact of one error type vs. the other would make us less confident in using this model to predict song success.

## Conclusion

**NEED CONCLUSION**

